Automatic Extraction of JavaScript Formal Semantics from ECMAScript Specifications

### 2019-08-07 icse/compiler.tex
More writing for algorithm compiler

The \( \ires \) function generator takes a parse tree from the token parser
and generates an \( \ires \) function. The given conversion rules describe
the mapping from each parsing rule into corresponding \( \ires \) components.
The \( \ires \) is an imperative style programming languages that we proposed
in order to represent each step of abstract algorithms in ECMAScript specifications.

% IR_ES 더 설명하기

@@

>>>>

We define \( \ires \) language to give semantics into syntactic parse tree of
abstract algorithm steps. It has the following language features:

\begin{itemize}
\item \textbf{Dynamic Typing} We decide to use dynamic typing for \( \ires \)
because each variables in abstract algorithms are not statically typed.
Thus, variables do not have its own static types while each value of \( \ires \)
has its dynamic type.

\item \textbf{Imparative Style} \( \ires \) represents algorithm steps
as statements. Each statement changes the current state
that consists of environments and a heap.

\item \textbf{Higher-order Functions with Restricted Scopes} \( \ires \) supports
lambda functions as values but it has only restricted scopes. In each function,
only global variables, parameter, and newly defined local variables are accessible.
It means that the function closure does not capture the the current environemnt.
We use such restricted scopes becuase it is enough to represent abstract algorithms.

\item \textbf{Primitive Values} \( \ires \) supports ECMAScript primitive values except
the symbols that is newly introduced after ECMAScript 6. Thus, only five ECMAScript primitives
are defined in \( \ires \); numbers, strings, booleans, undefined, and null.
Moreover, \( \ires \) also supports the unique \( \code{absent} \) value to represent
the absence of parameters. For example, the second parameter \( \code{PreferredType} \)
of \( \code{ToPrimitive} \) algorithm in Figure~\ref{fig:to-primitive} might be
not present in some function calls. Then, the parameter has the \( \code{absent} \) value
to represent its absence.

\item \textbf{Abstract Data Types} \( \ires \) only supports three abstract data types;
\( \code{Record} \) for mapping from some values into other value,
\( \code{List} \) for sequenital data types,
and \( \code{Singletone} \) for singleton data types.
They are stored in corresponding addresses in the heap structure.
For example, ECMAScript environment records are represented as \( \code{Record} \) data types
from String values for identifier names into addresses.
The addresses point to other \( \code{Record} \) data types that represent ECMAScript bindings.
We treat a ECMAScript constants such as \( \code{empty} \) and primitive symbols as singleton
data types in \( \ires \).
\end{itemize}

To generate the corresponding \( \ires \) function from syntactic parse tree,
we describe the mapping from each parsing rule into the corresponding \( \ires \) components.
For example, we describe the following mapping for the previous example parsing rules:
\begin{lstlisting}[style=myScalastyle]
// statements
val Stmt = "Let" ~ varT ~ "be" ~ Expr ~ "." ^^ { case _ ~ x ~ _ ~ y => ESLet(x, y) }
// expressions
val Expr = (
  // identifiers
  varT |
  // return if abrupt
  "?" ~ Expr ^^ { _ ~ e => ESReturnIfAbrupt(e) } |
  // function calls
  textT ~ "(" ~ repsep(varT, ",") ~ ")" ^^ { case x ~ _ ~ y ~ _ => ESCall(x, y) } |
  // lists
  "<<" ~ repsep(Expr, ",") ~ ">>" ^^ { case _ ~ x ~ _ => ESList(e) }
)
\end{lstlisting}
Finally, sub-step \( i \) in the \( \code{ToPrimitive} \) is converted
into the following instructions:
\begin{lstlisting}[style=myCorestyle]
let result = ? (Call exoticToPrim input (new [hint]))
\end{lstlisting}

@@

To make easy write conversion rules, we support \textit{rule generation assistant}.
It consists of several helper functions to show the specific statistics of
the not yet convertable abstract algorithm steps. It make easy not only construct
conversion reuls from the scratch but also update existing conversion rules
for updated specifications. We explain each helper function and how to utilize
it to construct conversion rules.

\subsubsection{Statement format suggestions}

\inred{TODO}

\subsubsection{Fault localizations based on \( \tstar \) tokens}

\inred{TODO}
